---
title: "Homework 7: Strings + Logical Vectors (Graphical Perception)"
author: "MICR 475/575"
format:
  html:
    toc: true
    embed-resources: true
execute:
  warning: false
  message: false
editor: visual
---

## Instructions

In this homework you will practice:

-   importing a Google Forms CSV export
-   working with **strings**
-   building and using **logical vectors**
-   making data tidy with `pivot_longer()`
-   making a simple **box plot**

You will work with a Google Forms CSV export from our in-class graphical perception experiment.

------------------------------------------------------------------------

## Setup

```{r}
library(tidyverse)

csv_path <- "../data/handb_ptest.csv"   # <-- update if needed
```

------------------------------------------------------------------------

# Q1. Import the data

**Task:** Read the CSV into an object called `raw`, then preview it.

```{r}
handb_ptest <-read.csv("data/handb_ptest.csv")
```

------------------------------------------------------------------------

## Q2. Tidy the data (Method A: pivot_longer first, then rename question type variable)

In the raw Google Forms export, the response columns have long names that include the full question text. Your goal is a tidy data frame with one row per response, and columns:

-   `timestamp`
-   `chart_type` (values like `T1`, `T2`, ..., `T9`)
-   `percent` (numeric)

**Task A:** Start from `raw`, `pivot_longer()` everything *except* `timestamp`.

**Task B:** Convert the `percent` column to a number.

**Task C:** Use **string operations** (and at least one **logical vector**) to convert the messy `chart_type` column into clean labels.

> Hint 1: `parse_number(chart_type)` can pull the number out of the string.\
> Hint 2: A logical vector can help you keep only the question numbers you want.

**Task D:** Make a **box plot** of `percent` by `chart_type` (no log transform).

```{r}
#pivot table longer
hpb_clean_b <- pivot_longer(handb_ptest,
      cols = c(2:19),
      names_to = c("question"),
      values_to = "score"
)

# get rid of %
df <- hpb_clean_b %>% mutate(
  score = parse_number(as.character(score)))

#remove NA
df <- df %>% filter(!is.na(score))

# change question names
df <- df %>% mutate(
  question = str_replace(question, "The.smaller.is.approximately.___..of.the.larger", "Q")
)

# make a box plot

library(ggplot2)

ggplot(data = df, aes(x = factor(question), y = score, fill = factor(question))) + 
  geom_boxplot(outlier.colour = "red", outlier.shape = 8) + # Customize outlier 
  theme_minimal() # Use a clean theme

```

------------------------------------------------------------------------

# Q3. Tidy the data (Method B: rename columns first, then pivot_longer)

This method uses the idea that (after `Timestamp`) the remaining columns are the 9 chart types, in order. The following steps provide one way this can be done - if you can get to the right answer using a different method, that's fine too!

**Task A:** Create a new data frame `df` that keeps `timestamp` plus the 9 response columns.

**Task B:** use the paste0() function to generate a vector of neat column headings 'T1', 'T2', etc. (keep `timestamp` as-is).

**Task C:** use the function set_names() with your new vector of column names to rename your columns

**Task D:** `pivot_longer()` to make a tidy data frame with `timestamp`, `chart_type`, `percent` (numeric).

**Task E:** Make the same **box plot** of `percent` by `chart_type`.

```{r}
#change column titles
colnames(handb_ptest) <- c("time", "q1a", "q1b", "q2a", "q2b", "q3a", "q3b", "q4a", "q4b", "q5a", "q5b", "q6a", "q6b", "q7a", "q7b", "q8a", "q8b", "q9a", "q9b")

#get rid of %
cols_to_clean <- c(3, 5, 7, 9, 11, 13, 15, 17, 19)

  # Apply gsub to remove all punctuation from the specified columns
  handb_ptest[cols_to_clean] <- lapply(handb_ptest[cols_to_clean], function(x) gsub("[[:punct:]]", "", x))

#pivot table
hpb_clean <- pivot_longer(handb_ptest,
  cols = starts_with("q"),
  names_to = c("question", ".value"),
  names_pattern = "(q\\d+)(.*)" # Capture 'Q1' and then the rest
)

#remove outliers
hpb_clean_2<- hpb_clean[-c(109, 47, 27), ]

#graph results
ggplot(data = hpb_clean_2, aes(x = factor(question), y = as.numeric(b), fill = factor(question))) + 
  geom_boxplot(outlier.colour = "red", outlier.shape = 8) + # Customize outlier 
  theme_minimal() # Use a clean theme
```

------------------------------------------------------------------------

# Q4. Do the two methods agree?

**Task A:** Check that the boxplots from Q2 and Q3 are the same. It's ok if they aren't - just please comment on why they disagree

**Task B:** In 3–6 sentences, comment on what the box plots show. Does it give the same gist that we would expect from Heer and Bostock's results? If not, why not?

```{r}
# YOUR CODE HERE
```

A: Yes, the outputs using methods A and B are the same (other than the axis labels)

B: Overall, the box blot is showing the spread of answers for each question around the mean of the answers. We can compare the spread of each question by comparing the size of each box. Here we see that questions 1, 3, and 8 had little spread, while 5, 6, and 9 had a lot of spread. So overall, as a class we are better at comparing the sizes of bars in comparison to other shapes and we aren't good at comparing when data is presented on separate scales (The ones in boxes) or as segmented horizontal lines.
